\texorpdfstring {Implementing Equational Constraints \\ in a Functional Language} {Implementing Equational Constraints in a Functional Language}

### 2012-11-09 papers/wlp2011/paper.tex
Typo + more explanation of LazyBind evaluation

In the next section, we review the source language Curry
and the features considered in this paper.
Section~\ref{sec:Compilation} recapitulates the implementation scheme
of KiCS2 originaly presented in \cite{BrasselHanusPeemoellerReck11}.
Sections~\ref{sec:Unification} and~\ref{sec:FuncPatterns}
discuss our new extensions to implement unification modulo
functional evaluation and functional patterns, respectively.
Benchmarks demonstrating the usefulness of these extensions
are presented in Sect.~\ref{sec:Benchmarks}
before we conclude in Sect.~\ref{sec:Conclusions}.

@@

Our proposed implementation of functional patterns in KiCS2
has a structure that is quite similar to that of equational constraints
with the exception that variables could be also bound
to unevaluated expressions. 
Only if such variables are later
accessed, the expressions they are bound to are evaluated.
This can be achieved by adding a further alternative
to the type of decisions:
\begin{haskell}
  data Decision = $\ldots$ | LazyBind [Constraint]
\end{haskell}
%
The implementation of the lazy unification operation \ccode{=:<=}
is almost identical to the strict unification operation \ccode{=:=}
as shown in Sect.~\ref{sec:Unification}.
The only difference is in the rules where a free variable occurs
in the left argument. All these rules are replaced by the single
rule
\begin{haskell}
  Choice (FreeID i) _ _ =:<= x
    = Guard [i :=: LazyBind (lazyBind i x)] Success
\end{haskell}
where the auxiliary operation \code{lazyBind}
implements the demand-driven evaluation of the right argument \code{x}:
\begin{haskell}
  lazyBind :: ID -> a -> [Constraint]
  lazyBind i True  = [i :=: ChooseLeft]
  lazyBind i False = [i :=: ChooseRight]
\end{haskell}
The use of the additional \code{LazyBind} constructor allows the argument 
\code{x} to be stored in a binding constraint without evaluation 
(due to the lazy evaluation strategy of the
target language Haskell).
However, it is evaluated by \code{lazyBind} 
as far as necessary when its binding
is required by another part of the computation.
Similarly to equational constraints,
lazy bindings are processed by a solver when values are extracted.
In particular, if a variable has more than one lazy binding constraint
(which is possible if a functional pattern evaluates to a non-linear term),
the corresponding expressions are evaluated and unified
according to the semantics of functional patterns
\cite{AntoyHanus05LOPSTR}.

>>>>

In the next section, we review the source language Curry
and the features considered in this paper.
Section~\ref{sec:Compilation} recapitulates the implementation scheme
of KiCS2 originally presented in \cite{BrasselHanusPeemoellerReck11}.
Sections~\ref{sec:Unification} and~\ref{sec:FuncPatterns}
discuss our new extensions to implement unification modulo
functional evaluation and functional patterns, respectively.
Benchmarks demonstrating the usefulness of these extensions
are presented in Sect.~\ref{sec:Benchmarks}
before we conclude in Sect.~\ref{sec:Conclusions}.

@@

Our proposed implementation of functional patterns in KiCS2
has a structure that is quite similar to that of equational constraints
with the exception that variables could be also bound
to unevaluated expressions. 
Only if such variables are later
accessed, the expressions they are bound to are evaluated.
This can be achieved by adding a further alternative
to the type of decisions:
\begin{haskell}
  data Decision = $\ldots$ | LazyBind [Constraint]
\end{haskell}
%
The implementation of the lazy unification operation \ccode{=:<=}
is almost identical to the strict unification operation \ccode{=:=}
as shown in Sect.~\ref{sec:Unification}.
The only difference is in the rules where a free variable occurs
in the left argument. All these rules are replaced by the single
rule
\begin{haskell}
  Choice (FreeID i) _ _ =:<= x
    = Guard [i :=: LazyBind (lazyBind i x)] Success
\end{haskell}
where the auxiliary operation \code{lazyBind}
implements the demand-driven evaluation of the right argument \code{x}:
\begin{haskell}
  lazyBind :: ID -> a -> [Constraint]
  lazyBind i True  = [i :=: ChooseLeft]
  lazyBind i False = [i :=: ChooseRight]
\end{haskell}
The use of the additional \code{LazyBind} constructor allows the argument 
\code{x} to be stored in a binding constraint without evaluation 
(due to the lazy evaluation strategy of the
target language Haskell).
However, it is evaluated by \code{lazyBind} to head normal form
when its binding is required by another part of the computation,
whereas the binding constraints for any sub-expression
are in turn lazily computed using `LazyBind`.

Similarly to equational constraints,
lazy bindings are processed by a solver when values are extracted.
In particular, if a variable has more than one lazy binding constraint
(which is possible if a functional pattern evaluates to a non-linear term),
the corresponding expressions are evaluated and unified
according to the semantics of functional patterns
\cite{AntoyHanus05LOPSTR}.
